---
title: "lab10"
author: "sl"
date: "`r Sys.Date()`"
output: github_document
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE) 
options(repos = c(CRAN = "http://cran.rstudio.com"))
```

## Data setup

```{r}
if (!require(RSQLite)) install.packages(c("RSQLite"))
if (!require(DBI)) install.packages(c("DBI"))

# install.packages(c("RSQLite", "DBI"))

library(RSQLite)
library(DBI)
```

```{r}
# Initialize a temporary in memory database
con <- dbConnect(SQLite(), ":memory:")

# Download tables
actor <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/actor.csv")
rental <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/rental.csv")
customer <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/customer.csv")
payment <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/payment_p2007_01.csv")

# Copy data.frames to database
dbWriteTable(con, "actor", actor)
dbWriteTable(con, "rental", rental)
dbWriteTable(con, "customer", customer)
dbWriteTable(con, "payment", payment)
dbListTables(con)
```

```{sql, connection = con, output.var = "x1"}
PRAGMA table_info(actor)
```

```{r}
x1
```

```{sql, connection = con}
PRAGMA table_info(actor)
```

## Exercise 1
Retrieve the actor ID, first name and last name for all actors using the actor table. Sort by last name and then by first name.

```{r}
dbGetQuery(con,
  "
  SELECT actor_id, first_name, last_name
  FROM actor
  ORDER by first_name, first_name
  LIMIT 10
  ")
```

## Use SQL directly

```{sql, connection = con}
SELECT actor_id, first_name, last_name
FROM actor
ORDER by first_name, first_name
LIMIT 10
```

## Exercise 2
Retrieve the actor ID, first name, and last name for actors whose last name equals ‘WILLIAMS’ or ‘DAVIS’.

```{r}
dbGetQuery(con,
  "
  SELECT actor_id, first_name, last_name
  FROM actor
  WHERE last_name IN ('WILLIAMS', 'DAVIS')
  ")
```

```{sql, connection = con}
SELECT actor_id, first_name, last_name
FROM actor
WHERE last_name IN ('WILLIAMS', 'DAVIS')
```

## Exercise 3
Write a query against the rental table that returns the IDs of the customers who rented a film on July 5, 2005 (use the rental.rental_date column, and you can use the date() function to ignore the time component). Include a single row for each distinct customer ID.

```{r}
dbGetQuery(con,
  "
  SELECT DISTINCT customer_id, rental_date
  FROM rental
  WHERE date(rental_date) = '2005-07-05'
  LIMIT 10
  ")
```

```{r}
#dbGetQuery(con,
#  "
#  SELECT customer_id, COUNT(*) AS N
#  FROM rental
#  WHERE date(rental_date) = '2005-07-05'
#  GROUP BY customer_id
#  ")
```

```{sql, connection = con}
SELECT DISTINCT customer_id, rental_date
FROM rental
WHERE date(rental_date) = '2005-07-05'
LIMIT 10
```

## Exercise 4
### Exercise 4.1
Construct a query that retrieve all rows from the payment table where the amount is either 1.99, 7.99, 9.99.

```{r}
dbGetQuery(con,
  "
  SELECT  *
  FROM payment
  WHERE amount IN (1.99, 7.99, 9.99)
  LIMIT 10
  ")
```

```{sql, connection = con}
SELECT  *
FROM payment
WHERE amount IN (1.99, 7.99, 9.99)
LIMIT 10
```

### Exercise 4.2
Construct a query that retrieve all rows from the payment table where the amount is greater then 5

```{r}
dbGetQuery(con,
  "
  SELECT  *
  FROM payment
  WHERE amount > 5
  LIMIT 10
  ")
```

```{sql, connection = con}
SELECT  *
FROM payment
WHERE amount > 5
LIMIT 10
```
### Exercise 4.3
Construct a query that retrives all rows from the payment table where the amount is greater then 5 and less then 8

```{r}
dbGetQuery(con,
  "
  SELECT  *
  FROM payment
  WHERE amount > 5 AND amount < 8
  LIMIT 10
  ")
```

```{sql, connection = con}
SELECT  *
FROM payment
WHERE amount > 5 AND amount < 8
LIMIT 10
```

## Exercise 5
Retrieve all the payment IDs and their amount from the customers whose last name is ‘DAVIS’.

```{r}
dbGetQuery(con,
  "
  SELECT c.customer_id, c.last_name, p.payment_id, p.amount
  FROM customer AS c INNER JOIN payment AS p
  ON c.customer_id = p.customer_id
  WHERE c.last_name = 'DAVIS'
  /*WHERE c.last_name IN ('DAVIS')*/
  ")
```

```{sql, connection = con}
SELECT c.customer_id, c.last_name, p.payment_id, p.amount
FROM customer AS c INNER JOIN payment AS p
ON c.customer_id = p.customer_id
WHERE c.last_name = 'DAVIS'
```

## Exercise 6
### Exercise 6.1
Use COUNT(*) to count the number of rows in rental

```{r}
dbGetQuery(con,
  "
  SELECT  COUNT(*) AS N
  FROM rental
  ")
```

```{sql, connection = con}
SELECT  COUNT(*) AS N
FROM rental
```

### Exercise 6.2
Use COUNT(*) and GROUP BY to count the number of rentals for each customer_id

```{r}
dbGetQuery(con,
  "
  SELECT customer_id, COUNT(*) AS N
  FROM rental
  GROUP BY customer_id
  LIMIT 10
  ")
```

```{sql, connection = con}
SELECT customer_id, COUNT(*) AS N
FROM rental
GROUP BY customer_id
LIMIT 10
```

### Exercise 6.3
Repeat the previous query and sort by the count in descending order

```{r}
dbGetQuery(con,
  "
  SELECT customer_id, COUNT(*) AS N
  FROM rental
  GROUP BY customer_id
  ORDER BY N DESC
  LIMIT 10
  ")
```

```{sql, connection = con}
SELECT customer_id, COUNT(*) AS N
FROM rental
GROUP BY customer_id
ORDER BY N DESC
LIMIT 10
```

### Exercise 6.4
Repeat the previous query but use HAVING to only keep the groups with 40 or more.

```{r}
dbGetQuery(con,
  "
  SELECT customer_id, COUNT(*) AS N
  FROM rental
  GROUP BY customer_id
  HAVING N >= 40
  ORDER BY N DESC
  LIMIT 10
  ")
```

```{sql, connection = con}
SELECT customer_id, COUNT(*) AS N
FROM rental
GROUP BY customer_id
HAVING N >= 40
ORDER BY N DESC
LIMIT 10
```

## Exercise 7
The following query calculates a number of summary statistics for the payment table using MAX, MIN, AVG and SUM

### Exercise 7.1
Modify the above query to do those calculations for each customer_id

```{r}
dbGetQuery(con,
  "
  SELECT customer_id,
    MAX(amount) AS max,
    MIN(amount) AS min,
    AVG(amount) AS avg,
    SUM(amount) AS sum
  FROM payment
  ")
```

```{sql, connection = con}
SELECT customer_id,
    MAX(amount) AS max,
    MIN(amount) AS min,
    AVG(amount) AS avg,
    SUM(amount) AS sum
FROM payment
```

### Exercise 7.2
Modify the above query to only keep the customer_ids that have more then 5 payments

```{r}

```

```{r}

```

## Cleanup
Run the following chunk to disconnect from the connection.
```{r}
# clean up
dbDisconnect(con)
```

