---
title: "lab05"
author: "sl"
date: "`r Sys.Date()`"
output: github_document
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(webshot)
library(lubridate)
library(tidyverse)
library(leaflet)
library(data.table)
library(dtplyr)
library(dplyr)
library(ggplot2)
```

```{r}
setwd("/Users/samuellu/Desktop/PM566/GitHub/pm566-fall2022-labs_Sam/lab05/")
```

## Step1: Read in the data
```{r read-data, cashe=TRUE}
if (!file.exists("../lab03/met_all.gz"))
  download.file(
    url = "https://raw.githubusercontent.com/USCbiostats/data-science-data/master/02_met/met_all.gz",
    destfile = "met_all.gz",
    method   = "libcurl",
    timeout  = 60
    )
met <- data.table::fread("../lab03/met_all.gz")
```

Read in station data

```{r stations, cashe=TRUE}
# Download the data
stations <- fread("ftp://ftp.ncdc.noaa.gov/pub/data/noaa/isd-history.csv")
stations[, USAF := as.integer(USAF)]

# Dealing with NAs and 999999
stations[, USAF   := fifelse(USAF == 999999, NA_integer_, USAF)]
stations[, CTRY   := fifelse(CTRY == "", NA_character_, CTRY)]
stations[, STATE  := fifelse(STATE == "", NA_character_, STATE)]

# Selecting the three relevant columns, and keeping unique records
stations <- unique(stations[, list(USAF, CTRY, STATE)])

# Dropping NAs
stations <- stations[!is.na(USAF)]

# Removing duplicates
stations[, n := 1:.N, by = .(USAF)]
stations <- stations[n == 1,][, n := NULL]
```

Remove temperatures less than -17C. Make sure there are no missing data in the key variables coded as 9999, 999, etc

```{r remove-low-temp}
met <- met[met$temp > -17][elev == 9999.0, elev:=NA]
```

Merge met data with stations.

```{r merge-datatables}
met <- 
  merge(
  # Data
  x     = met,      
  y     = stations, 
  # List of variables to match
  by.x  = "USAFID",
  by.y  = "USAF", 
  # Which obs to keep?
  all.x = TRUE,      
  all.y = FALSE
  ) 
```

## Question 1: Representative station for the US

Compute mean temp, wind.sp and atm.ps for each weather station, and pick the weather station with the average value closet to the median for the US.

```{r}
station_avg <-
  met[ , .(
    temp = mean(temp, na.rm = T),
    wind.sp = mean(wind.sp, na.rm = T),
    atm.press = mean(atm.press, na.rm = T)
  ), by = USAFID]
```

The above computes the mean by weather station.
Now let's compute the median value for each variable.

```{r}
stmeds <- station_avg[ , .(
  temp50 = median(temp, na.rm = T),
  windsp50 = median(wind.sp, na.rm = T),
  atmpress50 = median(atm.press, na.rm = T)
)]
```

A helpful function we might want to use 'which.min()'.

```{r}
station_avg[ ,
        temp_dist50 := abs(temp - stmeds$temp50)][order(temp_dist50)]

```

Let's use which.min

```{r}
station_avg[ which.min(temp_dist50)]
```

It matches the result above.








