---
title: "assignment03"
author: "sl"
date: "`r Sys.Date()`"
output: github_document
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rvest)
library(httr)
library(tidyverse)
library(stringr)
library(data.table)
library(xml2)
```

```{r}
setwd("/Users/samuellu/Desktop/PM566/GitHub/pm566-fall2022-labs_Sam/Assignment/assignment03/")
```

## Using the NCBI API, look for papers that show up under the term “sars-cov-2 trial vaccine.” Look for the data in the pubmed database, and then retrieve the details of the paper as shown in lab 7. How many papers were you able to find?

```{r}
# Downloading the website
website <- xml2::read_html("https://pubmed.ncbi.nlm.nih.gov/?term=sars-cov-2%20trial%20vaccine")

# Finding the counts
counts <- xml2::xml_find_first(website, "/html/body/main/div[9]/div[2]/div[2]/div[1]/div[1]")

# Turning it into text
counts <- as.character(counts)

# Extracting the data using regex
stringr::str_extract(counts, "[0-9,]+")
```

## Using the list of pubmed ids you retrieved, download each papers’ details using the query parameter rettype = abstract. If you get more than 250 ids, just keep the first 250.

```{r}
library(httr)
query_ids <- GET(
  url   = "https://eutils.ncbi.nlm.nih.gov/entrez/eutils/esearch.fcgi",
  query = list(
    db      = "pubmed",
    term = "sars-cov-2 trial vaccine.",
    rettype = "abstract",
    retmax  = 250)
)

# Extracting the content of the response of GET
ids <- httr::content(query_ids)
```

Get details about the articles

```{r}
# Turn the result into a character vector
ids <- as.character(ids)

# Find all the ids 
ids <- stringr::str_extract_all(ids, "<Id>[[:digit:]]+</Id>")[[1]]

# Remove all the leading and trailing <Id> </Id>. Make use of "|"
ids <- stringr::str_remove_all(ids, "</?Id>")

head(ids)
```

```{r}
publications <- GET(
  url   = "https://eutils.ncbi.nlm.nih.gov/entrez/eutils/efetch.fcgi",
  query = list(
    db  = "pubmed",
    id  = paste(ids,collapse = ","),
  retmax = 250,
  rettype = "abstract"
    )
)
# Extracting the content of the response of GET
publications <- httr::content(publications)
```

## As we did in lab 7. Create a dataset containing the following:

Pubmed ID number,
Title of the paper,
Name of the journal where it was published,
Publication date, and
Abstract of the paper (if any).

```{r}
pub_char_list <- xml2::xml_children(publications)
pub_char_list <- sapply(pub_char_list, as.character)
```

Now, extract the abstract and article title for each one of the elements of pub_char_list. You can either use sapply() as we just did, or simply take advantage of vectorization of stringr::str_extract

```{r}
abstracts <- str_extract(pub_char_list, "<Abstract>[[:print:][:space:]]+</Abstract>")
#abstracts[[1]]
abstracts <- str_remove_all(abstracts, "</?[[:alnum:]- =\"]+>") 
#abstracts[[1]]
abstracts <- str_replace_all(abstracts, "[[:space:]]+"," ")
abstracts[[1]]
```

Now get the titles:

```{r}
titles <- str_extract(pub_char_list, "<ArticleTitle>[[:print:][:space:]]+</ArticleTitle>")
#titles[[1]]
titles <- str_remove_all(titles, "</?[[:alnum:]- =\"]+>")
titles[[1]]
```

## Name of the journal where it was published

```{r}
names <- str_extract(pub_char_list, "<Title>[[:print:][:space:]]+</Title>")
#names[[1]]
names <- str_remove_all(names, "</?[[:alnum:]- =\"]+>")
names[[1]]
```

## Publication date

```{r}
years <- str_extract(pub_char_list, "<Year>[[:print:]]+</Year>")
#years[[1]]
years <- str_remove_all(years, "</?[[:alnum:]- =\"]+>")
years[[1]]
```

```{r}
months <- str_extract(pub_char_list, "<Month>[[:print:]]+</Month>")
#months[[1]]
months <- str_remove_all(months, "</?[[:alnum:]- =\"]+>")
months[[1]]
```

```{r}
days <- str_extract(pub_char_list, "<Day>[[:print:]]+</Day>")
#days[[1]]
days <- str_remove_all(days, "</?[[:alnum:]- =\"]+>")
days[[1]]
```
Finally the dataset:

```{r}
database <- data.frame(
  PubMedId = ids,
  Title    = titles,
  Journal  = names,
  Year     = years,
  month    = months,
  Day      = days,
  Abstract = abstracts
)
knitr::kable(database[1:8,], caption = "Some papers about sars-cov-2 trial vaccine.")
```

