---
title: "assignment03"
author: "sl"
date: "`r Sys.Date()`"
output: github_document
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rvest)
library(httr)
library(stringr)
library(data.table)
library(xml2)
```

```{r}
setwd("/Users/samuellu/Desktop/PM566/GitHub/pm566-fall2022-labs_Sam/Assignment/assignment03/")
```
# APIs

## Using the NCBI API, look for papers that show up under the term “sars-cov-2 trial vaccine.” Look for the data in the pubmed database, and then retrieve the details of the paper as shown in lab 7. How many papers were you able to find?

```{r}
# Downloading the website
website <- xml2::read_html("https://pubmed.ncbi.nlm.nih.gov/?term=sars-cov-2%20trial%20vaccine")

# Finding the counts
counts <- xml2::xml_find_first(website, "/html/body/main/div[9]/div[2]/div[2]/div[1]/div[1]")

# Turning it into text
counts <- as.character(counts)

# Extracting the data using regex
stringr::str_extract(counts, "[0-9,]+")
```

I was able to find 4,005 papers that show up under the term “sars-cov-2 trial vaccine.”

## Using the list of pubmed ids you retrieved, download each papers’ details using the query parameter rettype = abstract. If you get more than 250 ids, just keep the first 250.

```{r}

query_ids <- GET(
  url   = "https://eutils.ncbi.nlm.nih.gov/entrez/eutils/esearch.fcgi",
  query = list(
    db     = "pubmed", 
    term   = "sars-cov-2 trial vaccine", 
    retmax = 250
    )
)

# Extracting the content of the response of GET
ids <- httr::content(query_ids)
```

### Get PubMed ID number

```{r}
# Turn the result into a character vector
ids <- as.character(ids)

# Find all the ids 
ids <- stringr::str_extract_all(ids, "<Id>[[:digit:]]+</Id>")[[1]]

# Remove all the leading and trailing <Id> </Id>. Make use of "|"
ids <- stringr::str_remove_all(ids, "</?Id>")

head(ids)
```

```{r}
publications <- GET(
  url   = "https://eutils.ncbi.nlm.nih.gov/entrez/eutils/efetch.fcgi",
  query = list(
    db  = "pubmed",
    id  = paste(ids,collapse = ","),
  retmax = 250,
  rettype = "abstract"
    )
)
# Extracting the content of the response of GET
publications <- httr::content(publications)
```

## Create a dataset containing the following:

PubMed ID number, Title of the paper, Name of the journal where it was published, Publication date, and Abstract of the paper (if any).

```{r}
pub_char_list <- xml2::xml_children(publications)
pub_char_list <- sapply(pub_char_list, as.character)
```

### Get titles

```{r}
titles <- str_extract(pub_char_list, "<ArticleTitle>[[:print:][:space:]]+</ArticleTitle>")
#titles[[1]]
titles <- str_remove_all(titles, "</?[[:alnum:]- =\"]+>")
titles[[1]]
```

### Get name of the journal where it was published

```{r}
names <- str_extract(pub_char_list, "<Title>[[:print:][:space:]]+</Title>")
#names[[1]]
names <- str_remove_all(names, "</?[[:alnum:]- =\"]+>")
names[[1]]
```

### Get publication date

```{r}
years <- str_extract(pub_char_list, "<Year>[[:print:]]+</Year>")
#years[[1]]
years <- str_remove_all(years, "</?[[:alnum:]- =\"]+>")
years[[1]]
```

```{r}
months <- str_extract(pub_char_list, "<Month>[[:print:]]+</Month>")
#months[[1]]
months <- str_remove_all(months, "</?[[:alnum:]- =\"]+>")
months[[1]]
```

```{r}
days <- str_extract(pub_char_list, "<Day>[[:print:]]+</Day>")
#days[[1]]
days <- str_remove_all(days, "</?[[:alnum:]- =\"]+>")
days[[1]]
```

### Get abstracts

```{r}
abstracts <- str_extract(pub_char_list, "<Abstract>[[:print:][:space:]]+</Abstract>")
#abstracts[[1]]
abstracts <- str_remove_all(abstracts, "</?[[:alnum:]- =\"]+>") 
#abstracts[[1]]
abstracts <- str_replace_all(abstracts, "[[:space:]]+"," ")
abstracts[[1]]
```

### Create my dataset:

```{r}
database <- data.frame(
  PubMedId = ids,
  Title    = titles,
  Journal  = names,
  Year     = years,
  Month    = months,
  Day      = days,
  Abstract = abstracts
)
knitr::kable(database[1:3,], align = "lllllll", caption = "Some papers about sars-cov-2 trial vaccine")

```

I was able to get Year, Month, and Day. However, I still have to find out a way to merge them.

# Text Mining

```{r install-libraries}
library(tidytext)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(forcats)
```

## Read in the data

```{r read-data, cashe=TRUE}
if (!file.exists("pubmed.csv"))
  download.file(
    url = "https://raw.githubusercontent.com/USCbiostats/data-science-data/master/03_pubmed/pubmed.csv",
    destfile = "pubmed.csv",
    method   = "libcurl",
    timeout  = 60
    )
pm <- read.csv("pubmed.csv")
#dim(pm)
str(pm)
pm <- as_tibble(pm)
pm
```

## 1. Tokenize the abstracts and count the number of each token. Do you see anything interesting? Does removing stop words change what tokens appear as the most frequent? What are the 5 most common tokens for each search term after removing stopwords?

```{r}
pm %>%
  unnest_tokens(token, abstract) %>%
  count(token) %>%
  top_n(10, n) %>%
  ggplot(aes(n, fct_reorder(token, n))) +
  geom_col()
```

I can only find "covid" and "19" which are interesting because there are lots of stop words. It would be nice to remove stop words.

```{r}
pm %>%
  unnest_tokens(token, abstract) %>%
  anti_join(stop_words, by = c("token" = "word")) %>%
  count(token) %>%
  top_n(10, n) %>%
  ggplot(aes(n, fct_reorder(token, n))) +
  geom_col()
```

After removing stop words, the word "covid" became the most frequent word and we can also find there are other interesting words appeared to the top 10 list.

```{r top5, cache = TRUE}
pm %>%
  unnest_tokens(token, abstract) %>%
  group_by(term) %>%
  count(token, sort = TRUE) %>%
  filter(!(token %in% stop_words$word) & !grepl(pattern = "^[0-9]+$", x = token)) %>%
  top_n(5, n) %>%
  arrange(term, desc(n)) %>%
knitr::kable()
```

After removing stop words, I created a list with the 5 most common tokens for each search term. We can understand that each search term contains its own related words from this list.

2. Tokenize the abstracts into bigrams. Find the 10 most common bigram and visualize them with ggplot2.

```{r}
pm %>%
  unnest_ngrams(ngram, abstract, n = 2) %>%
  count(ngram, sort = TRUE) %>%
  top_n(10, n) %>%
  ggplot(aes(n, fct_reorder(ngram, n))) +
  geom_col()
```

Although we may tokenize the abstracts into bigrams, I was wondering whether I can remove those stop words in bigrams.

3. Calculate the TF-IDF value for each word-search term combination. (here you want the search term to be the “document”) What are the 5 tokens from each search term with the highest TF-IDF value? How are the results different from the answers you got in question 1?

```{r}
pm %>%
  unnest_tokens(token, abstract) %>%
  group_by(term) %>%
  count(token, sort = TRUE) %>%
  filter(!(token %in% stop_words$word)) %>%
  top_n(5, n) %>%
  bind_tf_idf(token, term, n) %>%
  arrange(term, desc(tf_idf)) %>%
knitr::kable()
```

This list shows the top 5 tokens from each search term with the highest TF-IDF value. The results are more relevant to their terms and it can can be easier to target their main tokens with one glance than the method from question 1.

