---
title: "assignment01"
author: "sl"
date: "`r Sys.Date()`"
output: github_document
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r library}
library(webshot)
library(lubridate)
library(tidyverse)
library(data.table)
library(leaflet)
```

## Step 1. Conduct EDA Checklist items 2-4

### Read in the data - 2004 and 2019
```{r}
setwd("/Users/samuellu/Desktop/PM566/GitHub/pm566-fall2022-labs_Sam/Assignment/assignment01/")
data_2004 <- data.table::fread("ad_viz_plotval_data_2004.csv")
data_2019 <- data.table::fread("ad_viz_plotval_data_2019.csv")
```
### Check the dimensions and headers and footers of the data
```{r}
dim(data_2004)
head(data_2004)
tail(data_2004)
```
```{r}
dim(data_2019)
head(data_2019)
tail(data_2019)
```
### Check the variable types in the data
```{r}
str(data_2004)
```
```{r}
str(data_2019)
```

### Check for any data issues, particularly in the key variable we are analyzing
```{r}
#table(data_2004$Date)
#table(data_2004$`Site Name`)
#table(data_2004$CBSA_NAME)
#table(data_2004$STATE)
table(data_2004$COUNTY)
```
```{r}
#table(data_2019$Date)
#table(data_2019$`Site Name`)
#table(data_2019$CBSA_NAME)
#table(data_2019$STATE)
table(data_2019$COUNTY)
```

```{r}
#summary(data_2004$`Daily Mean PM2.5 Concentration`)
summary(data_2004$CBSA_CODE) #1253 NA's
#summary(data_2004$COUNTY_CODE)
#summary(data_2004$SITE_LATITUDE)
#summary(data_2004$SITE_LONGITUDE)
```
```{r}
#summary(data_2019$`Daily Mean PM2.5 Concentration`)
summary(data_2019$CBSA_CODE) #4181 NA's
#summary(data_2019$COUNTY_CODE)
#summary(data_2019$SITE_LATITUDE)
#summary(data_2019$SITE_LONGITUDE)
```

There are 1253 NAs in CBSA_CODE of data_2004 and 4181 NAs in CBSA_CODE of data_2019 but it is fine.

## Step 2. Combine the two years of data into one data frame. Use the Date variable to create a new column for year, which will serve as an identifier. 
```{r}
data_2004_2019 <- rbind(data_2004, data_2019) 
pm <- mutate(data_2004_2019, year = factor(rep(c(2004, 2019), c(nrow(data_2004), nrow(data_2019))))) %>%
        rename(PM2.5 = `Daily Mean PM2.5 Concentration`)
```

Create a factor variable indicating which year the data comes from. I also rename the Daily Mean PM2.5 Concentration variable to a more sensible PM2.5

```{r}
with(pm, tapply(PM2.5, year, summary))
```

From the summary of 2004 and 2019 appear there are some negative values of PM2.5.

### Change the names of the key variables so that they are easier to refer to in your code.
```{r}
colnames(pm)[19] <- "lat"
colnames(pm)[20] <- "lon"
```
```{r}
colnames(data_2004_2019)[5] <- "PM2.5"
colnames(data_2004_2019)[19] <- "lat"
colnames(data_2004_2019)[20] <- "lon"
```
```{r}
colnames(data_2004)[5] <- "PM2.5"
colnames(data_2004)[19] <- "lat"
colnames(data_2004)[20] <- "lon"

colnames(data_2019)[5] <- "PM2.5"
colnames(data_2019)[19] <- "lat"
colnames(data_2019)[20] <- "lon"
```

## Step 3. Create a basic map in leaflet() that shows the locations of the sites (make sure to use different colors for each year). 

```{r}
leaflet() %>%
  addProviderTiles('OpenStreetMap') %>% 
  addCircles(data = unique(pm[year == "2004"]),
             lat=~lat,lng=~lon, 
             color = "blue",
             opacity = 0.1, fillOpacity = 1, radius = 500) %>% 
  addCircles(data = unique(pm[year == "2019"]),
             lat=~lat,lng=~lon, 
             color = "red",
             opacity = 0.1, fillOpacity = 1, radius = 100)
```
###  Summarize the spatial distribution of the monitoring sites.

It seems that there are more monitoring sites in the bit cities like San Jose, LA, San Francisco and San Diego.

## Step 4. Check for any missing or implausible values of PM2.5 in the combined dataset
```{r}
summary(pm$PM2.5)
```

```{r}
mean(is.na(pm$PM2.5))
```

```{r}
nrow(pm[PM2.5 <0])
```

There is no missing values of PM2.5 in the combined dataset. However, it is implausible to have negative value of PM2.5.

```{r}
filter(pm, year == "2004") %>% summarize(negative = mean(PM2.5 < 0, na.rm = TRUE))
```
```{r}
filter(pm, year == "2019") %>% summarize(negative = mean(PM2.5 < 0, na.rm = TRUE))
```
However, there is a relatively small proportion of values that are negative.

### Explore the proportions of each and provide a summary of any temporal patterns you see in these observations.

```{r}
length(unique(pm[year ==2004]$`Site ID`))
```
```{r}
length(unique(pm[year ==2019]$`Site ID`))
```

```{r}
mean(pm[year==2004]$PM2.5)
```
```{r}
mean(pm[year==2019]$PM2.5)
```

Monitoring sites were getting more and more. The total mean pm2.5 level was lower in 2019 than in 2004.

## Step 5. Explore the main question of interest at three different spatial levels(state, county, site in Los Angeles)
### Create exploratory plots (e.g.boxplots, histograms, line plots)
### summary statistics that best suit each level of data
### Be sure to write up explanations of what you observe in these data.

First, I create 2004 and 2019 datasets with Date, STATE, COUNTY, Site Name, pm2.5, lat, and lon by Site ID.

```{r avg}
data_2004_avg <- data_2004[, .(
  Date = Date,
  STATE = STATE,
  COUNTY = COUNTY,
  `Site Name` =`Site Name`,
  PM2.5 = mean(PM2.5, na.rm = T),
  lat = mean(lat, na.rm = T), 
  lon = mean(lon, na.rm = T)
), by="Site ID"]

data_2004_avg$year <- "2004"

data_2019_avg <- data_2019[, .(
  Date = Date,
  STATE = STATE,
  COUNTY = COUNTY,
  `Site Name` =`Site Name`,
  PM2.5 = mean(PM2.5, na.rm = T),
  lat = mean(lat, na.rm = T), 
  lon = mean(lon, na.rm = T)
), by="Site ID"]

data_2019_avg$year <- "2019"
```

Second, I merged two datasets.

```{r}
pm_avg <- rbind(data_2004_avg, data_2019_avg)
```
Add one column of region.
```{r}
pm_avg[, region := fifelse(lon >= -98 & lat > 39.71, "NE",
                fifelse(lon < -98 & lat > 39.71, "NW",
                fifelse(lon < -98 & lat <= 39.71, "SW","SE")))
   ]
table(pm_avg$region)
```

Add on column of pm_health (pm2.5 <= 12 is labeled as health).
```{r}
pm_avg[, pm_health := fifelse(PM2.5 <= 12, "healthy",
                         fifelse(PM2.5 > 12 & PM2.5 < 35, "normal", "unhealthy"))]

table(pm_avg$pm_health)
```

### Box plots

```{r}
#set.seed(2015)
#idx <- sample(nrow(pm), 1000)
qplot(year , log2(PM2.5), data = pm_avg, geom = "boxplot")
```

From the raw boxplot, it seems that on average, the levels of PM in 2019 are lower than they were in 2004.

```{r}
pm_avg[STATE == "California"] %>% 
  ggplot() + 
  geom_boxplot(mapping = aes(x = year, y = PM2.5, color=pm_health, fill = pm_health)) + 
  facet_wrap(~ region, nrow = 1)
```

In California, southwestern region pm2.5 level is higher than northwestern region.

```{r}
pm_avg[COUNTY == "Los Angeles"] %>% 
  ggplot() + 
  geom_boxplot(mapping = aes(x = year, y = PM2.5, color=pm_health, fill = pm_health)) + 
  facet_wrap(~ region, nrow = 1)
```

In Los Angeles, pm2.5 level is higher in 2004 than in 2019.

```{r}
pm_avg[`Site Name` == "Los Angeles-North Main Street"] %>% 
  ggplot() + 
  geom_boxplot(mapping = aes(x = year, y = PM2.5, color=pm_health, fill = pm_health)) + 
  facet_wrap(~ region, nrow = 1)
```

In the Los Angeles-North Main Street, pm2.5 level turns into healthy in 2019.

### Violin plots
```{r}
ggplot(data = pm_avg[STATE == "California"], aes(x=year, y=PM2.5, color=pm_health, fill = pm_health)) +
 geom_violin() +
 geom_point() +
 facet_wrap(~ region, nrow = 1)
```

California pm2.5 level is lower in 2019 than in 2004 in both northwestern and southwestern.

```{r}
ggplot(data = pm_avg[COUNTY == "Los Angeles"], aes(x=year, y=PM2.5, color=pm_health, fill = pm_health)) +
 geom_violin() +
 geom_point() +
 facet_wrap(~ region, nrow = 1)
```

In Los Angeles, most pm2.5 levels is healthier in 2019 than in 2004.

```{r}
ggplot(data = pm_avg[`Site Name` == "Los Angeles-North Main Street"], aes(x=year, y=PM2.5, color=pm_health, fill = pm_health)) +
 geom_violin() +
 geom_point()
```

On main street in LA, its pm2.5 level used to be higher in 2004 but it's within the healthy range in 2019.

### Histograms
```{r}
ggplot(data = pm_avg[STATE == "California"], aes(x=PM2.5, color=year, fill = year)) +
  geom_histogram(fill="white", alpha=0.5, position="identity") +
   facet_wrap(~ region, nrow = 1)
```

In this histogram, there are more data in southwestern region than in northwestern region.

```{r}
ggplot(data = pm_avg[COUNTY == "Los Angeles"], aes(x=PM2.5, color=year, fill = year)) +
  geom_histogram(fill="white", alpha=0.5, position="identity") +
  facet_wrap(~ region, nrow = 1)
```

In Los Angeles, there are more counts in the range of 15 to 20 in 2004 than in 2019.

```{r}
ggplot(data = pm_avg[`Site Name` == "Los Angeles-North Main Street"], aes(x=PM2.5, color=year, fill = year)) +
  geom_histogram(fill="white", alpha=0.5, position="identity") +
  facet_wrap(~ region, nrow = 1)
```

On the North Main Street, its used to get pm2.5 levels around 20 in 2004. However, it is lower in 2019.

### examining changes in PM at the county level. This analysis falls somewhere in between looking at the entire State all at once and looking at an individual monitor.

What we do here is calculate the mean of PM for each county in 2004 and 2019

```{r}
county_pm <- group_by(pm, year, COUNTY) %>% summarize(PM2.5 = mean(PM2.5, na.rm = TRUE))
head(county_pm)
```

```{r}
qplot(years, PM2.5, data = mutate(county_pm, years = as.numeric(as.character(year))), 
      color = factor(COUNTY), 
      geom = c("point", "line"))
```

It shows that many counties have decreased the average pm2.5 levels from 2004 to 2019 but a few states actually increased their levels.


## Whether daily concentrations of PM2.5 have decreased in California over the last 15 years (from 2004 to 2019)

From above data and plots, I would say that the daily concentrations of p2.5 have decreased in California over the last 15 years.


