---
title: "mid-term"
author: "sl"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r library}
library(lubridate)
library(tidyverse)
library(leaflet)
library(data.table)
library(dtplyr)
library(dplyr)
library(ggplot2)
library(stringr)
```

```{r}
setwd("/Users/samuellu/Desktop/PM566/GitHub/pm566-fall2022-labs_Sam/mid_term/")
```

# Introduction (provide background on your dataset and formulated question)
These two datasets are about "diagnosed diabetes among adults aged >=18 years" and "Obesity among adults aged >=18 years" in 2017 from the CDC. They include estimates for the 500 largest US cities and approximately 28,000 census tracts within these cities. With these two datasets, first question is that I want to know each city highest, lowest rates of diabetes and obesity and their total length. Second, I would like to know whether there is a correlation between diabetes and obesity. Last, I want to compare different regions diabetes_percentage and obesity_percentage with histograms and boxplots.

# Methods (include how and where the data were acquired, how you cleaned and wrangled the data, what tools you used for data exploration)

## Read in the data by API

I used API method to obtain my datasets from CDC. First, you have to create an account with password. Then, you have to apply for a free app token. Last, copy your API Endpoint. Here are my datasets links:
`https://chronicdata.cdc.gov/500-Cities-Places/500-Cities-Obesity-among-adults-aged-18-years/bjvu-3y7d`
`https://chronicdata.cdc.gov/500-Cities-Places/500-Cities-Diagnosed-diabetes-among-adults-aged-18/cn78-b9bj`

```{r API_dia, cashe=TRUE}
#install.packages("RSocrata")
library("RSocrata")

dia <- read.socrata(
  "https://chronicdata.cdc.gov/resource/cn78-b9bj.json?year=2017",
  app_token = "bEkVW73ASzmTkZ9riAtf2YS5c",
  email     = "clu74108@usc.edu",
  password  = "Samuelsunny0325!"
)
dia <- as.data.table(dia)
write.csv(dia,"dia.csv", row.names = F)
```

```{r API_obe, cashe=TRUE}
obe <- read.socrata(
  "https://chronicdata.cdc.gov/resource/bjvu-3y7d.json?year=2017",
  app_token = "bEkVW73ASzmTkZ9riAtf2YS5c",
  email     = "clu74108@usc.edu",
  password  = "Samuelsunny0325!"
)
obe <- as.data.table(obe)
write.csv(obe,"obe.csv", row.names = F)
```

```{r}
#dia <- fread("dia.csv")
#dia <- as.data.table(dia)
#obe <- fread("obe.csv")
#obe <- as.data.table(obe)
```

Afer getting my datasets, I check their dim and str to see which columns I am going to keep.

```{r}
dim(dia)
dim(obe)
```

```{r}
str(dia)
str(obe)
```

## Select columns

I select data_value, populationcount, stateabbr, statedesc, cityname, geolocation.latitude, and geolocation.longitude total 7 columns.

```{r}
dia_mini <- dia[, c(2, 3, 14, 17, 19, 20, 23)]
obe_mini <- obe[, c(2, 3, 14, 17, 19, 20, 23)]
```

## Change column names

I change my column names into the way that is easy to understand.

```{r}
colnames(dia_mini)[1] <- "diabetes_percentage"
colnames(dia_mini)[2] <- "dia_PopulationCount"
colnames(dia_mini)[3] <- "state"
colnames(dia_mini)[6] <- "lat"
colnames(dia_mini)[7] <- "lon"

colnames(obe_mini)[1] <- "obesity_percentage"
colnames(obe_mini)[2] <- "obe_PopulationCount"
colnames(obe_mini)[3] <- "state"
colnames(obe_mini)[6] <- "lat"
colnames(obe_mini)[7] <- "lon"
```

## Merge two datasets

```{r merge-datatables}
merged <- 
  merge(
  # Data
  x     = dia_mini,      
  y     = obe_mini, 
  # List of variables to match
 by = c("state", "statedesc", "cityname", "lat", "lon"),
  # keep everything!
  all.x = TRUE     
  ) 

dim(merged)
```

I merge two datasets by state, statedesc, cityname, lat, and lon. However, my row number increased so I have to remove duplicates.

```{r}
merged[, n := 1:.N, by = .(state, statedesc, cityname, lat, lon)]
merged <- merged[n == 1,][, n := NULL]

length(unique(merged$lat))
```

After removing duplicates, my rows shrink from 30,008 to 28,505.

## Convert columns which are character variables into numeric variables

```{r}
str(merged)
```

```{r}
merged$lat <- as.numeric(merged$lat)
merged$lon <- as.numeric(merged$lon)
merged$diabetes_percentage <- as.numeric(merged$diabetes_percentage)
merged$dia_PopulationCount <- as.numeric(merged$dia_PopulationCount)
merged$obesity_percentage <- as.numeric(merged$obesity_percentage)
merged$obe_PopulationCount <- as.numeric(merged$obe_PopulationCount)

str(merged)
```

## Check NAs values in my dataset

```{r}
summary(merged$diabetes_percentage)
summary(merged$obesity_percentage)
```

```{r}
mean(is.na(merged$diabetes_percentage))
mean(is.na(merged$obesity_percentage))
```

However, there are only 2.7% NAs in my dataset, which is not significant. Therefore, I'm going to replace NA values with mean.

```{r}
merged[, diabetes_percentage := fcoalesce(diabetes_percentage, mean(diabetes_percentage, na.rm = TRUE))]
merged[, obesity_percentage := fcoalesce(obesity_percentage, mean(obesity_percentage, na.rm = TRUE))]
```

```{r}
summary(merged$diabetes_percentage)
summary(merged$obesity_percentage)
```

# Preliminary Results (provide summary statistics in tabular form and publication-quality figures, take a look at the kable function from knitr to write nice tables in Rmarkdown)

## Show min, max, median and length of diabetes_percentage and obesity_percentage columns in different cities.

```{r}
status_merged <- merged[, .(
  diabetes_min = min(diabetes_percentage, na.rm = T),
  diabetes_max = max(diabetes_percentage, na.rm = T),
  diabetes_median = median(diabetes_percentage, na.rm = T),
  diabetes_length = length(diabetes_percentage),
  obesity_min = min(obesity_percentage, na.rm = T),
  obesity_max = max(obesity_percentage, na.rm = T),
  obesity_median = median(obesity_percentage, na.rm = T),
  obesity_length = length(obesity_percentage)
), by="cityname"][order(-diabetes_median)]

status_merged
```

```{r}
merged[, .(
  diabetes_min = min(diabetes_percentage, na.rm = T),
  diabetes_max = max(diabetes_percentage, na.rm = T),
  diabetes_median = median(diabetes_percentage, na.rm = T),
  diabetes_length = length(diabetes_percentage),
  obesity_min = min(obesity_percentage, na.rm = T),
  obesity_max = max(obesity_percentage, na.rm = T),
  obesity_median = median(obesity_percentage, na.rm = T),
  obesity_length = length(obesity_percentage)
), by="cityname"][order(diabetes_median)]
```

```{r}
merged[, .(
  diabetes_min = min(diabetes_percentage, na.rm = T),
  diabetes_max = max(diabetes_percentage, na.rm = T),
  diabetes_median = median(diabetes_percentage, na.rm = T),
  diabetes_length = length(diabetes_percentage),
  obesity_min = min(obesity_percentage, na.rm = T),
  obesity_max = max(obesity_percentage, na.rm = T),
  obesity_median = median(obesity_percentage, na.rm = T),
  obesity_length = length(obesity_percentage)
), by="cityname"][order(-obesity_median)]
```

```{r}
merged[, .(
  diabetes_min = min(diabetes_percentage, na.rm = T),
  diabetes_max = max(diabetes_percentage, na.rm = T),
  diabetes_median = median(diabetes_percentage, na.rm = T),
  diabetes_length = length(diabetes_percentage),
  obesity_min = min(obesity_percentage, na.rm = T),
  obesity_max = max(obesity_percentage, na.rm = T),
  obesity_median = median(obesity_percentage, na.rm = T),
  obesity_length = length(obesity_percentage)
), by="cityname"][order(obesity_median)]
```

Among all cities, Gary, in NE region, has both the highest mean of diabetes_percentage(24.05%) and obesity_percentage(50.30%). On the other hand, Boulder, in NW region, has both the lowest mean of diabetes_percentage(5.00%) and obesity_percentage(15.70%).

## Scatter plot

From the scatter plot by states, we can see that there is a positive correlation between obesity and diabetes rates in different states.

```{r scatter}
ggplot(merged, aes(x =obesity_percentage, y=diabetes_percentage, color = state)) + 
  geom_point(size=0.1, position="jitter") +
  geom_smooth(size=0.2, method=lm, se=FALSE, fullrange=TRUE, color = "black")
```

## Add regions: Northeast, Northwest, Southwest, and Southeast.

I create a new column contain four different regions

```{r}
merged[, region := fifelse(lon >= -98 & lat > 39.71, "NE",
                fifelse(lon < -98 & lat > 39.71, "NW",
                fifelse(lon < -98 & lat <= 39.71, "SW","SE")))
   ]
table(merged$region)
```

```{r scatter_region}
ggplot(merged[!is.na(region)], aes(x =obesity_percentage, y=diabetes_percentage, color = region)) + 
  geom_point(size=0.1, position="jitter") +
  geom_smooth(size=0.2, method=lm, se=FALSE, fullrange=TRUE, color = "black") +
  scale_color_brewer(palette="Set2") +
  scale_fill_brewer() +
  facet_wrap(~region, nrow = 2)
```
In different regions, there are all in a positive correlation between obesity_percentage and diabetes_percentage. Apart from the result, we can also see that the data points in the Northwest are fewer than in other regions.

## Use leaflet to visulize diabetes_percentage and obesity_percentage

First, create a color palette so we can see the severity of each place.

```{r}
pal_dia <- colorNumeric(c('darkblue','goldenrod','darkred'), domain=merged$diabetes_percentage)
```

```{r}
#label, color
leaflet() %>%
  addProviderTiles('OpenStreetMap') %>% 
  addCircles(data = merged,
             lat=~lat,lng=~lon,
             label = ~paste0(round(diabetes_percentage,2)), color = ~ pal_dia(diabetes_percentage),
             opacity = 0.5, fillOpacity = 1, radius = 50) %>%
  # And a pretty legend
  addLegend('bottomleft', pal=pal_dia, values=merged$diabetes_percentage,
             title='diabetes%', opacity=1)
```

Then, I would like to take a look at LA where I live in.

```{r}
#label, color
leaflet() %>%
  addProviderTiles('OpenStreetMap') %>% 
  addCircles(data = merged[merged$cityname == "Los Angeles"],
             lat=~lat,lng=~lon,
             label = ~paste0(round(diabetes_percentage,2)), color = ~ pal_dia(diabetes_percentage),
             opacity = 0.5, fillOpacity = 1, radius = 50) %>%
  # And a pretty legend
  addLegend('bottomleft', pal=pal_dia, values=merged$diabetes_percentage,
             title='diabetes%', opacity=1)
```

We can see that cities closer to Downtown LA have higher rates of diabetes in orange color.

```{r}
pal_obe <- colorNumeric(c('darkblue','goldenrod','darkred'), domain=merged$obesity_percentage)
```

```{r}
#label, color
leaflet() %>%
  addProviderTiles('OpenStreetMap') %>% 
  addCircles(data = merged[merged$cityname == "Los Angeles"],
             lat=~lat,lng=~lon,
             label = ~paste0(round(obesity_percentage,2)), color = ~ pal_obe(obesity_percentage),
             opacity = 0.5, fillOpacity = 1, radius = 50) %>%
  # And a pretty legend
  addLegend('bottomleft', pal=pal_obe, values=merged$obesity_percentage,
             title='obesity%', opacity=1)
```

Similar to the result of diabetes_percentage,the region near downtown LA turn orange into red.

## Histogram

I would like to compare diabetes_percentage in different regions.

```{r his_dia}
ggplot(data = merged[!is.na(diabetes_percentage)][!is.na(region)], aes(diabetes_percentage, color=region, fill = region)) + 
  geom_histogram(fill="white", alpha=0.5) +
  scale_color_brewer(palette="Set2")
```
From this histogram, we can see that NE region has higher counts of diabetes_percentage. However, it might be affected by different number and size of cities in different regions.

```{r}
table(merged$region)
```

```{r his_obe}
ggplot(data = merged[!is.na(obesity_percentage)][!is.na(region)], aes(obesity_percentage, color=region, fill = region)) + 
  geom_histogram(fill="white", alpha=0.5) +
  scale_color_brewer(palette="Set2")
```

we can see that NE region has higher counts of obesity_percentage as well. However, it might be also affected by different number and size of cities in different regions.

## Check diabetes_percentage's min, max, and median between different regions.

```{r}
merged[!is.na(diabetes_percentage)][!is.na(region)] %>%
  ggplot() + 
    stat_summary(mapping = aes(x = region, y = diabetes_percentage),
    fun.min = min,
    fun.max = max,
    fun = median)
```

There is a highest diabetes_percentage almost 40% in NE region.

```{r}
merged[!is.na(obesity_percentage)][!is.na(region)] %>%
  ggplot() + 
    stat_summary(mapping = aes(x = region, y = obesity_percentage),
    fun.min = min,
    fun.max = max,
    fun = median)
```
In NE region, there is a max obesity_percentage occurred. However, SE region has the highest median of obesity_percentage.

## Boxplot

```{r box_dia}
ggplot(merged[!is.na(diabetes_percentage)][!is.na(region)], aes(x=region, y=diabetes_percentage, color=region)) +
  geom_boxplot() +
  scale_color_brewer(palette="Set2")
```

```{r box_obe}
ggplot(merged[!is.na(obesity_percentage)][!is.na(region)], aes(x=region, y=obesity_percentage, color=region)) +
  geom_boxplot() +
  scale_color_brewer(palette="Set2")
```

And from the boxplot, NW region has the lowest median in diabetes_percentage and obesity_percentage.

# Conclusion

## Question 1: Which city has the highest median rates of diabetes and obesity and which city lowest rates of diabetes and obesity?

Among all cities, Gary, in NE region, has the highest median rates of diabetes(24.05%) and he highest median rates of  obesity(50.30%). On the other hand, Boulder, in NW region, has the lowest median rates of diabetes(5.00%) and the lowest median rates of obesity(15.70%).

## Question 2: What is the correlation between diabetes and obesity in different region?

We can see that there is a positive correlation between obesity and diabetes rates from the scatter plot by states, so as the scatter plot by regions. Apart from the result, we can also see that the data points in the Northwest are fewer than in other regions.

## Question 3: Compare different regions diabetes_percentage and obesity_percentage with different plots(such as histograms ,boxplots, or leaflet).

### Leaflet
From the leaflet, cities closer to Downtown LA have higher rates of diabetes in orange color. Similar to the result of diabetes percentage, the region near downtown LA shows orange color dots as well.

### Histogram
From histograms, the NE region has higher counts of diabetes_percentage. We can also see that the NE region has higher counts of obesity_percentage as well. However, all results from histograms might be affected by different numbers and sizes of cities in different regions.

### Boxplot
There is the highest diabetes_percentage in the NE region(almost 40%). However, the SE region has the highest median of obesity_percentage. The NW region has both the lowest median in diabetes_percentage and obesity_percentage.